Recursive Solutions to First-Order Model Counting

### 2023-01-05 doc/paper/paper.tex
An implementation of the algorithm that constructs the definitions of functions from an FCG

@@

However, none of the publicly available exact (W)FOMC algorithms can efficiently
compute functions as simple as a factorial.\footnote{The problem of computing
  the factorial can be described using two variables and counting quantifiers,
  so it is known to be liftable in principle
  \citep{DBLP:journals/jair/Kuzelka21}, but an algorithm that can lift such an
  instance in practice does not exist.} We claim that this shortcoming is due to
the inability of these algorithms to construct recursive solutions. The topic of
recursion in the context of WFOMC has been studied before but in limited ways.
\Citet{DBLP:conf/ilp/BarvinekB0ZK21} use WFOMC to generate numerical data that
is then used to conjecture recurrence relations that explain that data.
\Citet{DBLP:conf/nips/Broeck11} introduced the idea of \emph{domain recursion}.
Intuitively, domain recursion partitions a domain of size $n$ into a single
explicitly named constant and the remaining domain of size $n-1$. However, many
stringent conditions are enforced to ensure that the search for a tractable
solution always terminates.

@@

\subsection{Generalised Domain Recursion}\label{sec:dr}

@@

\subsection{Constraint Removal}\label{sec:cr}

@@

\section{Empirical Results}\label{sec:results} % Newly Domain-Liftable Formulas

\begin{table*}[t]
  \centering
  \begin{tabular}{cccccc}
    \toprule
    \multicolumn{3}{c}{Function Class} & \multicolumn{3}{c}{Asymptotic Complexity of Counting} \\
    Partial & Endo- & Class & Best Known & With \textsc{ForcLift} & With \textsc{Crane} \\
    \midrule
    \rowcolor{gray!10}\cmark/\xmark & \cmark/\xmark & Functions & $\log m$ & $m$ & $m$ \\
    \xmark & \xmark & \multirow{4}{*}{Surjections} & $n \log m$ & $m^{3}+n^{3}$ & $m^{3}+n^{3}$ \\
    \xmark & \cmark & & $m \log m$ & $m^{3}$ & $m^{3}$ \\
    \cmark & \xmark & & \multicolumn{3}{c}{Same as injections from $\Delta$ to $\Gamma$} \\
    \cmark & \cmark & & \multicolumn{3}{c}{Same as endo-injections} \\
    \rowcolor{gray!10}\xmark & \xmark & & $m$ & --- & $mn$ \\
    \rowcolor{gray!10}\xmark & \cmark & & $m$ & --- & $m^3$ \\
    \rowcolor{gray!10}\cmark & \xmark & & $\min\{\, m, n \,\}^2$ & --- & $mn$ \\
    \rowcolor{gray!10}\cmark & \cmark & \multirow{-4}{*}{Injections} & $m^2$ & --- & --- \\
    \xmark & \xmark & \multirow{3}{*}{Bijections} & $m$ & --- & $m$ \\
    \xmark & \cmark & & \multicolumn{3}{c}{\multirow{2}{*}{Same as (partial) (endo-)injections}} \\
    \cmark & \cmark/\xmark & & \multicolumn{3}{c}{} \\
    \bottomrule
  \end{tabular}
  \caption{The worst-case complexity of counting various types of functions.
    Here, $m$ is the size of domain $\Gamma$, and $n$ is the size of domain
    $\Delta$. All asymptotic complexities are in $\Theta(\cdot)$. A dash means
    that no complete solution was found.}\label{tbl:results}
\end{table*}
% TODO: It is a standard to discuss a few lifted theories like the smokers and
% friends example. Can you also include some empirical results on such theories
% as well to appeal to the sanity of the reader that it actually works for
% problems in frameworks such as Markov log networks

>>>>

% TODO: when describing experimental results, say that a solution for the last
% (unsolved) instance can be found, but is too complicated to evaluate by hand.
% Add that the linear order axiom is also liftable. Cite the same paper in the
% introduction as well.

@@

However, none of the publicly available exact (W)FOMC algorithms can efficiently
compute functions as simple as a factorial.\footnote{There is previous work
  \citep{DBLP:journals/jair/Kuzelka21,DBLP:conf/ijcai/Broeck16} showing how FOMC
  can compute factorials in principle, but an algorithm that can lift such an
  instance in practice does not exist.} We claim that this shortcoming is due to
the inability of these algorithms to construct recursive solutions. The topic of
recursion in the context of WFOMC has been studied before but in limited ways.
\Citet{DBLP:conf/ilp/BarvinekB0ZK21} use WFOMC to generate numerical data that
is then used to conjecture recurrence relations that explain that data.
\Citet{DBLP:conf/nips/Broeck11} introduced the idea of \emph{domain recursion}.
Intuitively, domain recursion partitions a domain of size $n$ into a single
explicitly named constant and the remaining domain of size $n-1$. However, many
stringent conditions are enforced to ensure that the search for a tractable
solution always terminates.

@@

\subsection{Generalised Domain Recursion}\label{sec:dr}
% TODO: the thesis has some improvements to this section

@@

\subsection{Constraint Removal}\label{sec:cr}
% TODO: the thesis has an improved version of this section

@@

\section{Empirical Results}\label{sec:results} % Newly Domain-Liftable Formulas
% TODO: the thesis has an improved version of this section
% TODO: add the sample solutions as an appendix (they're already in the thesis)

\begin{table*}[t]
  \centering
  \begin{tabular}{cccccc}
    \toprule
    \multicolumn{3}{c}{Function Class} & \multicolumn{3}{c}{Asymptotic Complexity of Counting} \\
    Partial & Endo- & Class & Best Known & With \textsc{ForcLift} & With \textsc{Crane} \\
    \midrule
    \rowcolor{gray!10}\cmark/\xmark & \cmark/\xmark & Functions & $\log m$ & $m$ & $m$ \\
    \xmark & \xmark & \multirow{4}{*}{Surjections} & $n \log m$ & $m^{3}+n^{3}$ & $m^{3}+n^{3}$ \\
    \xmark & \cmark & & $m \log m$ & $m^{3}$ & $m^{3}$ \\
    \cmark & \xmark & & \multicolumn{3}{c}{Same as injections from $\Delta$ to $\Gamma$} \\
    \cmark & \cmark & & \multicolumn{3}{c}{Same as endo-injections} \\
    \rowcolor{gray!10}\xmark & \xmark & & $m$ & --- & $mn$ \\
    \rowcolor{gray!10}\xmark & \cmark & & $m$ & --- & $m^3$ \\
    \rowcolor{gray!10}\cmark & \xmark & & $\min\{\, m, n \,\}^2$ & --- & $mn$ \\
    \rowcolor{gray!10}\cmark & \cmark & \multirow{-4}{*}{Injections} & $m^2$ & --- & --- \\
    \xmark & \xmark & \multirow{3}{*}{Bijections} & $m$ & --- & $m$ \\
    \xmark & \cmark & & \multicolumn{3}{c}{\multirow{2}{*}{Same as (partial) (endo-)injections}} \\
    \cmark & \cmark/\xmark & & \multicolumn{3}{c}{} \\
    \bottomrule
  \end{tabular}
  \caption{The worst-case complexity of counting various types of functions.
    Here, $m$ is the size of domain $\Gamma$, and $n$ is the size of domain
    $\Delta$. All asymptotic complexities are in $\Theta(\cdot)$. A dash means
    that the algorithm was not able to find a solution.}\label{tbl:results}
\end{table*}
% TODO: It is a standard to discuss a few lifted theories like the smokers and
% friends example. Can you also include some empirical results on such theories
% as well to appeal to the sanity of the reader that it actually works for
% problems in frameworks such as Markov log networks
